The semantic drift of quotations in blogspace: a case study in short-term cultural evolution

### 2016-08-01 brains-conclusion.tex
manuscript revised

We aimed to connect the field of cultural evolution with psycholinguistics by asking if cultural attractors appear in a corpus of online news-related quotes which are gradually transformed by low-level biases.
The data set we used imposed a few constraints on our analysis:
first, it was necessary to infer source-destination links, an operation made more reliable when restricting the scope of transformations to very simple cases, which we did by focusing on word substitutions.
Second, contrary to laboratory experiments which produce data made of many repeated measures on a small number of cases (\hbox{e.g.} a given list of words), we have a great number of different cases (one case per cluster in which substitutions are found, \hbox{i.e.} 698 cases), with very few measures on each of them (average 9, median 5).
This rendered the prediction of individual words impractical: if we cannot compute a percentage of explained data for a given case, any approximate prediction will be heavily underestimated.
This last factor, added to the potential for variation of external conditions when authors wrote the quotes, led us to use word features to analyze the transformations by aggregating over individual cases.

@@

We do not actually observe quotes converging on a global scale towards attractors in their various dimensions.
Indeed the limits of the data set don't allow us to infer chains of substitutions, and substitutions themselves are not the only type of transformation at work in the data set.
Nonetheless, these findings (1) bring light to this simple type of transformation, and (2) are consistent with known psycholinguistic effects, with the hypothesis of cultural attractors in representations from everyday life, and with the lineage specificity discussed in the iterated learning literature~\citep{claidiere_cultural_2014,cornish_systems_2013}.
They are obtained by successfully applying knowledge from cognitive science to real-life complex data, a task that remains a challenge in the study of cultural evolution.
We believe that applying such data mining tools to manage the complexity of real-life data is a promising approach for the joint analysis of cognitive science and culture.

In the simple case presented here, however, much remains to be explored.
Since it is clear that observing cognitive biases in such data is now possible, questions addressed in controlled laboratory situations could be opened by further research.
One question concerns the influence of the context surrounding a quote, be it in terms of other quotes preceding it temporally or of text surrounding it in a post.
A first step could be the application of results from \citet{zaromb_temporal_2006} who have shown, in the simpler task of recall of random word lists, that the source of prior-list intrusions can be predicted based on the associations those preceding lists have formed:
in our case, a substitution could be triggered and directed by associations formed by preceding context.
A further step would be to follow \citet{cornish_systems_2013} have shown about reciprocal influences between context and transformations (in their case, with transmission chains of artificial content).
Indeed substitutions, and more generally all transformations, also participate in creating the context for later quotes.
One can ask, therefore, what are the reciprocal effects between, on one side, the corpus-level evolution of quotes through iterated transformations, and on the other side, a gradual change in the properties of transformations operated because of the evolution of surrounding context.
Such interactions have been shown to underlie the lineage specificity observed in transmission chains~\citep{claidiere_cultural_2014}.
Exploring how similar loop interactions happen real-life data could indeed be the next step in understanding the evolution of both cultural content and the ways in which it is transformed.
In our particular case, such insight could shed some light on how the feature attractors explored in this paper actually emerge.

@@

\newtext{The theory of Epidemiology of Representations proposes a unifying framework for the study of cultural evolution.
One of its core claims, the existence of cultural attractors, has been both a challenge to test empirically and a fruitful line to pursue in the study of cultural evolution.}
We aimed to contribute to \newtext{testing this hypothesis} by studying a simple everyday-life task where individuals are implicitly trying to reproduce quotations.
To some extent, our work amounts to \newtext{an out-of-laboratory} experiment where we \newtext{examine the influence} of \newtext{well-known word features} on the accuracy of the reproduction of short sentences.
\newtext{Our analysis of substitutions shows that words are attracted, in each dimension, to a feature-specific value.
Furthermore, the features' known effects in psycholinguistic experiments are reflected in the biases of these attraction points, meaning that the evolution of such quotations can be partially explained by known low-level cognitive biases.
We believe that such an approach, combining psycholinguistic knowledge and data mining tools, can be fruitfully developed to improve the study of cultural attractors and explore the reciprocal influences of cognition and culture.}

>>>>

We \tb{initially} aimed to connect the field of cultural evolution with psycholinguistics by asking if cultural attractors appear in a corpus of online news-related quotes which are gradually transformed by low-level biases.
The data set we used imposed a few constraints on our analysis:
first, it was necessary to infer source-destination links, an operation made more reliable when restricting the scope of transformations to very simple cases, which we did by focusing on word substitutions.
Second, contrary to laboratory experiments which produce data made of many repeated measures on a small number of cases (\hbox{e.g.} a given list of words), we have a great number of different cases (one case per cluster in which substitutions are found, \hbox{i.e.} 698 cases), with very few measures on each of them (average 9, median 5).
This rendered the prediction of individual words impractical: if we cannot compute a percentage of explained data for a given case, any approximate prediction will be heavily underestimated.
This last factor, added to the potential for variation of external conditions when authors wrote the quotes, led us to use word features to analyze the transformations by aggregating over individual cases.

@@

We do not actually observe quotes converging on a global scale towards attractors in their various dimensions.
Indeed the limits of the data set don't allow us to infer chains of substitutions, and substitutions themselves are not the only type of transformation at work in the data set.
Nonetheless, these findings (1) bring light to this simple type of transformation, and (2) are consistent with known psycholinguistic effects, with the hypothesis of cultural attractors in representations from everyday life, and with the lineage specificity discussed in the iterated learning literature~\citep{claidiere_cultural_2014,cornish_systems_2013}.
They are obtained by successfully applying knowledge from cognitive science to real-life complex data, a task that remains a challenge in the study of cultural evolution.
\tb{More broadly,} we believe that applying such data mining tools to manage the complexity of real-life data is a promising approach for the joint analysis of cognitive science and culture.

In the simple case presented here, however, much remains to be explored.
Since it is clear that observing cognitive biases in such data is now possible, questions addressed in controlled laboratory situations could be opened by further research.
One question concerns the influence of the context surrounding a quote, be it in terms of other quotes preceding it temporally or of text surrounding it in a post.
A first step could be the application of results from \citet{zaromb_temporal_2006} who have shown, in the simpler task of recall of random word lists, that the source of prior-list intrusions can be predicted based on the associations those preceding lists have formed:
in our case, a substitution could be triggered and directed by associations formed by preceding context.
A further step would be to follow \citet{cornish_systems_2013} have shown about reciprocal influences between context and transformations (in their case, with transmission chains of artificial content).
Indeed substitutions, and more generally all transformations, also participate in creating the context for later quotes.
One can ask, therefore, what are the reciprocal effects between, on one side, the corpus-level evolution of quotes through iterated transformations, and on the other side, a gradual change in the properties of transformations operated because of the evolution of surrounding context.
Such interactions have been shown to underlie the lineage specificity observed in transmission chains~\citep{claidiere_cultural_2014}.
Exploring how similar loop interactions happen in real-life data could indeed be the next step in understanding the evolution of both cultural content and the ways in which it is transformed.
In our particular case, such insight could shed some light on how the feature attractors explored in this paper actually emerge.

@@

\newtext{The theory of Epidemiology of Representations proposes a unifying framework for the study of cultural evolution.
One of its core claims, the existence of cultural attractors, has been both a challenge to test empirically and a fruitful line to pursue in the study of cultural evolution.}
We aimed to contribute to \newtext{testing this hypothesis} by studying a simple everyday-life task where individuals are implicitly trying to reproduce quotations.
To some extent, our work amounts to \newtext{an out-of-laboratory} experiment where we \newtext{examine the influence} of \newtext{well-known word features} on the accuracy of the reproduction of short sentences.
\newtext{Our analysis of substitutions shows that words are attracted, in each dimension, to feature-specific values.
Furthermore, the features' known effects in psycholinguistic experiments are reflected in the biases of these attraction points, meaning that the evolution of such quotations can be partially explained by known low-level cognitive biases.
We believe that such an approach, combining psycholinguistic knowledge and data mining tools, can be fruitfully developed to improve the study of cultural attractors and explore the reciprocal influences of cognition and culture.}

### 2016-08-01 brains-protocol.tex
manuscript revised

While most of the substitutions we obtain with this procedure replace a word with another semantically related word, the two are rarely direct synonyms:
only a third of all substitutions travel less than 3 hops on the hyponym-hypernym network defined by WordNet (direct synonyms count as 0~hops on this network), meaning that at least two thirds involve non-synonyms.\footnote{
\newtext{A similar behavior is observed on the FA network, where about 104 clusters have substitutions traveling only 1 hop, 110 traveling 2 hops, 137 traveling 3, 72 traveling 4, and 13 traveling 5.
}}
Compared to the word it replaces, the new word usually achieves a similar meaning in the context of its sentence while still slightly changing the implications or the attitude expressed by the author.
The following examples illustrate this behavior:

@@

@@

\begin{figure}[h]
    \centering
    \includegraphics[width=\linewidth]{images/computed-figures/all-susceptibilities-pos.png}
    \caption{\newtext{\textbf{Substitution susceptibility for POS categories:} categories are simplified from the TreeTagger tag set: \emph{J} means adjective, \emph{N}~noun, \emph{R}~adverb, \emph{V}~verb, and \emph{Stopword-like} gathers together all the categories partially or completely affected by the stopword filter (see main text for details).
    The top panel shows the actual $s_{POS}$ and $s_{POS}^0$ counts, the bottom panel shows $\sigma_{POS}$, the ratio of the two.
    Confidence intervals are computed with the Goodman method for multinomial proportions.}
    }
    \label{fig:pos-susceptibilities}
\end{figure}

@@

We say that a word is \emph{substitutable} if it appears in a quote which undergoes a substitution, whether the substitution operates on that word or on another one.
\newtext{For a given group of words $g$, say all nouns, or all words in a particular range of values for a feature (\hbox{e.g.} words 2 to 4 letters long), susceptibility is computed as the ratio of $s_g$, the number of times words of that group are substituted, to $s_g^0$, the number of times words of that group would be substituted if substitutions fell randomly on substitutable words.
That is:}\footnote{
\newtext{$s_g^0$ is computed by summing, over all quotes undergoing a substitution, the ratio of the number of words in a quote that are from group $g$ to the number of words in the same quote that could have been the substitution's target.
If, for instance, half the content words of a given quote are nouns, such a quote contributes .5 to the total $s_{nouns}^0$.} \tb{Further, t}o avoid possible autocorrelation effects due to substitutions belonging to the same cluster (which are likely not statistically independent and may lead to overly optimistic confidence intervals), we scale $s_g$ and $s_g^0$ to count one for each cluster.
That is, each individual substitution contributes to $s_g$ or $s_g^0$ by only 1 over the number of substitutions in its cluster.}
}
$$\sigma_g = \frac{s_g}{s_g^0}$$

@@

We also introduce \newtext{two} null \newtext{hypotheses}, $\mathcal{H}_0$ \newtext{and $\mathcal{H}_{00}$}, to compare the actual variation of a word's feature to expected variations \newtext{under unbiased transformations}.
\newtext{$\mathcal{H}_0$ models the situation where} the arrival word $\warrival$ \newtext{is} randomly chosen from the whole pool of words available in the data set for that feature.\footnote{
For instance, when considering the feature ``Clustering coefficient'', the arrival word is randomly chosen among words present in the data set of FA norms.
}
In this case, since $\phi(\warrival)$ becomes a constant value in the above averaging (by definition $\warrival$ does not depend on $\wstart$ anymore), the baseline variation under $\mathcal{H}_0$ may be rewritten as:
$$\nu_{\phi}^0 (f) = \left<\phi\right>$$
\newtext{$\mathcal{H}_{00}$ models the situation where the arrival word $\warrival$ is chosen \emph{among immediate synonyms of the start word} $\wstart$, that is an arrival word chosen among semantically plausible though still random words.
In this case $\nu_{\phi}^{00}$ does depend on $f$:\footnote{
\newtext{The actual implementation has an additional level of averaging since WordNet, used to get a word's synonyms, defines several meanings for a single given word.
Therefore:
$$\nu_{\phi}^{00} (f) = \left< \left< \left< \phi(\warrival) \right>_{\warrival \in syn(m)} \right>_{m \in meanings(\wstart)} \right>_{\left\lbrace \wstart | \phi(\wstart) = f \right\rbrace}$$}
}
$$\nu_{\phi}^{00} (f) = \left< \left< \phi(\warrival) \right>_{\warrival \in syn(\wstart)} \right>_{\left\lbrace \wstart | \phi(\wstart) = f \right\rbrace}$$}

>>>>

\tb{Substitutions usually} replace a word with another semantically related word: the new word achieves a similar meaning in the context of its sentence, while still slightly changing the implications or the attitude expressed by the author. The following examples illustrate this \tb{phenomenon}:

@@

\tb{However, the substituted and subtituting words are rarely direct synonyms:} only a third of all substitutions travel less than 3 hops on the hyponym-hypernym network defined by WordNet (direct synonyms count as 0~hops on this network), meaning that at least two thirds involve non-synonyms.\footnote{
\newtext{A similar \tb{phenomenon} is observed on the FA network, where about 104 clusters have substitutions traveling only 1 hop, 110 traveling 2 hops, 137 traveling 3, 72 traveling 4, and 13 traveling 5.
}}

%Compared to the word it replaces, the new word usually achieves a similar meaning in the context of its sentence while still slightly changing the implications or the attitude expressed by the author.
%The following examples illustrate this behavior:

%\begin{APAitemize}
%	\item ``This is \{socialism~$\rightarrow$~welfare\} for the rich'',
%	\item {[}The{]} ``perverse logic of \{clashes~$\rightarrow$~confrontation\} and violence'',
%	\item ``This \{crisis~$\rightarrow$~problem\} did not develop overnight and it will not be solved overnight''.
%\end{APAitemize}

@@

\begin{figure}[h]
    \centering
    \includegraphics[width=\linewidth]{images/computed-figures/all-susceptibilities-pos.png}
    \caption{\newtext{\textbf{\tb{POS-related results:}} categories are simplified from the TreeTagger tag set: \emph{J} means adjective, \emph{N}~noun, \emph{R}~adverb, \emph{V}~verb, and \emph{Stopword-like} gathers together all the categories partially or completely affected by the stopword filter (see main text for details).
    The top panel shows the actual $s_{POS}$ and $s_{POS}^0$ counts. The bottom panel shows \tb{the substitution susceptibility $\sigma_{POS}$, which is the ratio between the two previous counts}.
    Confidence intervals are computed with the Goodman method for multinomial proportions.}
    }
    \label{fig:pos-susceptibilities}
\end{figure}

@@

We say that a word is \emph{substitutable} if it appears in a quote which undergoes a substitution, whether the substitution operates on that word or on another one.
\newtext{For a given group of words $g$, say all nouns, or all words in a particular range of values for a feature (\hbox{e.g.} words 2 to 4 letters long), susceptibility is computed as the ratio of $s_g$, the number of times words of that group are substituted, to $s_g^0$, the number of times words of that group would be substituted if substitutions fell randomly on substitutable words.
That is:}\footnote{
\newtext{$s_g^0$ is computed by summing, over all quotes undergoing a substitution, the ratio of the number of words in a quote that are from group $g$ to the number of words in the same quote that could have been the substitution's target.
If, for instance, half the content words of a given quote are nouns, such a quote contributes .5 to the total $s_{nouns}^0$.} \tb{Further, t}o avoid possible autocorrelation effects due to substitutions belonging to the same cluster (which are likely not statistically independent and may lead to overly optimistic confidence intervals), we scale $s_g$ and $s_g^0$ to count one for each cluster.
That is, each individual substitution contributes to $s_g$ or $s_g^0$ by only 1 over the number of substitutions in its cluster.}

$$\sigma_g = \frac{s_g}{s_g^0}$$

@@

We also introduce \newtext{two} null \newtext{hypotheses}, $\mathcal{H}_0$ \newtext{and $\mathcal{H}_{00}$}, to compare the actual variation of a word's feature to expected variations \newtext{under unbiased transformations}.
\newtext{$\mathcal{H}_0$ models the situation where} the arrival word $\warrival$ \newtext{is} randomly chosen from the whole pool of words available in the data set for that feature.\footnote{
For instance, when considering the feature ``Clustering coefficient'', the arrival word is randomly chosen among words present in the data set of FA norms.
}
In this case, since $\phi(\warrival)$ becomes a constant value in the above averaging (by definition $\warrival$ does not depend on $\wstart$ anymore), the baseline variation under $\mathcal{H}_0$ may be rewritten as:
$$\nu_{\phi}^0 (f) = \left<\phi\right>$$
\newtext{$\mathcal{H}_{00}$ models the situation where the arrival word $\warrival$ is chosen \emph{among immediate synonyms of the start word} $\wstart$, \hbox{i.e.} an arrival word chosen among semantically plausible though still random words.
In this case $\nu_{\phi}^{00}$ does depend on $f$:\footnote{
\newtext{The actual implementation has an additional level of averaging since WordNet, used to get a word's synonyms, defines several meanings for a single given word.
Therefore:
$$\nu_{\phi}^{00} (f) = \left< \left< \left< \phi(\warrival) \right>_{\warrival \in syn(m)} \right>_{m \in meanings(\wstart)} \right>_{\left\lbrace \wstart | \phi(\wstart) = f \right\rbrace}$$}
}
$$\nu_{\phi}^{00} (f) = \left< \left< \phi(\warrival) \right>_{\warrival \in syn(\wstart)} \right>_{\left\lbrace \wstart | \phi(\wstart) = f \right\rbrace}$$}

### 2016-08-01 brainscopycut.tex
beginning of introduction - end of conclusion
manuscript revised

% To make comments, corrections, notes
\newcommand{\tb}[1]{\textcolor{blue}{#1}}
\newcommand{\tg}[1]{\textcolor{MidnightBlue}{#1}}
\newcommand{\newtext}[1]{\tg{#1}}
\newenvironment{new}{\par\color{MidnightBlue}}{\par}
\newcommand{\add}[1]{\tb{ADD: #1}}
\newcommand{\ra}[1]{$\rightarrow$#1}
\newcommand{\replace}[2]{\sout{#1} \tb{#2}}
\newcommand{\rk}[1]{\tb{{\footnotesize {\bf[\emph{#1}]}}}}
\newcommand{\marginnote}[1]{\marginpar{{\parbox{1.\linewidth}{\hrulefill\\\tb{\scriptsize \sf #1}}}}}
\newcommand{\cam}[1]{\tb{\small{[{\bf C:} }#1{]}}}
\newcommand{\seb}[1]{\tb{\small{[{\bf S:} #1{]}}}}
\newcommand{\todo}[1]{\tb{#1}}
\definecolor{lightblue}{rgb}{.85,.85,1}
\definecolor{lightgreen}{rgb}{.85,1,.85}

@@

\title{The semantic drift of quotations in blogspace: a case study in short-term cultural evolution}
\shorttitle{The semantic drift of quotations in blogspace}
\date{}
\twoauthors{Sébastien Lerique}{Camille Roth}
\twoaffiliations{Centre d'Analyse et de Mathématique Sociales, UMR 8557 CNRS/EHESS\\190 av. de France, F-75013 Paris\\and Centre Marc Bloch Berlin, UMIFRE 14 CNRS/MAEE/HU\\Friedrichstr. 191, D-10117 Berlin}{CNRS\\Centre Marc Bloch Berlin, UMIFRE 14 CNRS/MAEE/HU\\Friedrichstr. 191, D-10117 Berlin}
\abstract{
\newtext{We present a case study which connects psycholinguistics with the field of cultural evolution, in order to test for the existence of cultural attractors in the evolution of quotations online.
Such attractors have been suggested as a useful tool for the study of cultural evolution and its reciprocal influences with cognition, but their existence has been hard to test.
Our study focuses on how quotations are transformed when they are copied from blog to blog or media website:
by coding words with a number of well-studied lexical features, we show that the way words are substituted in quotations is consistent (1) with the hypothesis of cultural attractors, and (2) with known effects of the word features.
In particular, words known to be harder to recall in lists have a higher tendency to be substituted, and words easier to recall are produced instead.
Our results support the hypothesis that cultural attractors can arise from the combination of individual cognitive biases in interpretation and production of representations.}
}
\authornote{Correspondence should be directed to lerique@cmb.hu-berlin.de and roth@cmb.hu-berlin.de}
\keywords{\newtext{word recall}; recollection bias; semantic network; cultural evolution; cultural attraction; data mining; big data; \emph{in vivo} psycholinguistics}
%\note{\TB{Note XXXX (version? date?)}}

>>>>

% To make comments, corrections, notes
\newcommand{\tb}[1]{\textcolor{blue}{#1}}
\newcommand{\tbn}[1]{\tb{\sout{#1}}}
\newcommand{\tg}[1]{\textcolor{MidnightBlue}{#1}}
\newcommand{\newtext}[1]{\tg{#1}}
\newenvironment{new}{\par\color{MidnightBlue}}{\par}
\newcommand{\add}[1]{\tb{ADD: #1}}
\newcommand{\ra}[1]{$\rightarrow$#1}
\newcommand{\replace}[2]{\sout{#1} \tb{#2}}
\newcommand{\rk}[1]{\tb{{\footnotesize {\bf[\emph{#1}]}}}}
\newcommand{\marginnote}[1]{\marginpar{{\parbox{1.\linewidth}{\hrulefill\\\tb{\scriptsize \sf #1}}}}}
\newcommand{\cam}[1]{\tb{\small{[{\bf C:} }#1{]}}}
\newcommand{\seb}[1]{\tb{\small{[{\bf S:} #1{]}}}}
\newcommand{\todo}[1]{\tb{#1}}
\definecolor{lightblue}{rgb}{.85,.85,1}
\definecolor{lightgreen}{rgb}{.85,1,.85}

@@

\title{The semantic drift of quotations in blogspace: a case study in short-term cultural evolution}
\shorttitle{The semantic drift of quotations in blogspace}
\date{}
\twoauthors{Sébastien Lerique}{Camille Roth}
\twoaffiliations{Centre d'Analyse et de Mathématique Sociales, UMR 8557 CNRS/EHESS\\190 av. de France, F-75013 Paris\\and Centre Marc Bloch Berlin, UMIFRE 14 CNRS/MAEE/HU\\Friedrichstr. 191, D-10117 Berlin}{CNRS\\Centre Marc Bloch Berlin, UMIFRE 14 CNRS/MAEE/HU\\Friedrichstr. 191, D-10117 Berlin}
\abstract{
\newtext{We present a case study which connects psycholinguistics with the field of cultural evolution, in order to test for the existence of cultural attractors in the evolution of quotations.
Such attractors have been \tb{proposed} as \tb{a concept useful for} understanding cultural evolution %and its reciprocal influences 
\tb{in relation} with \tb{individual} cognition, but their existence has been hard to test.
\tb{We empirically} focus on \tb{the online transformation of} quotations when they are copied from blog to blog or media website: by coding words with a number of well-studied lexical features, we show that the way words are substituted in quotations is consistent (1) with the hypothesis of cultural attractors, and (2) with known effects of the word features.
In particular, words known to be harder to recall in lists have a higher tendency to be substituted, and words easier to recall are produced instead.
Our results support the hypothesis that cultural attractors can \tb{result} from the combination of individual cognitive biases in the interpretation and reproduction of representations.}
}
\authornote{Correspondence should be directed to lerique@cmb.hu-berlin.de and roth@cmb.hu-berlin.de}
\keywords{\newtext{word recall}; recollection bias; semantic network; cultural evolution; cultural attraction; data mining; big data; \emph{in vivo} psycholinguistics}
%\note{\TB{Note XXXX (version? date?)}}
